---
title: "PA1_template"
output: html_document
---

Peer Assignment 1
=================

set the libraries required for the code

```{r}
library(ggplot2)
library(lattice)
library(plyr)
```


Loading and preprocessing the data. Also aggregate the steps per day

```{r}
Activity_Csv_File<-"./activity.csv"
Activity<-read.csv(Activity_Csv_File,header=TRUE)
Steps_perday<-aggregate(Activity$steps,by=list(Activity$date),FUN=sum,na.rm=TRUE)
names(Steps_perday)<-c("date","steps")

```


####What is mean total number of steps taken per day?

Histogram showing total number of steps taken each day

```{r}
th<-theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank(),text=element_text(size=12))
p<-ggplot(Steps_perday,aes(x=steps,fill=..count..))+geom_histogram()
gt<-ggtitle("Histogram of steps taken each day")
myplot<-p+gt+th+ylab("Count of Steps")
myplot
```

Summary statistics below showing the mean and median total number of steps  taken per day

```{r}
summary(Steps_perday$steps)
```

####What is the average daily activity pattern?

Aggregate the data by Interval to calculate avg steps by 5-minute interval

```{r}
AvgStep_Interval<-aggregate(Activity$steps,by=list(Activity$interval),FUN=mean,na.rm=TRUE)
names(AvgStep_Interval)<-c("interval","Avg_Steps")
```

Time series plot showing average steps by 5-minute interval

```{r}
p<-ggplot(AvgStep_Interval,aes(x=interval,y=Avg_Steps))+geom_line()
gt<-ggtitle("Average Steps by Interval")
myplot<-p+gt+th+ylab("Average Steps")+xlab("Interval")
myplot
```

Number of missing rows
 
```{r} 
sum(is.na(Activity$steps))
```


####Imputing missing values

Impute the missing values by mean of 5-minute interval

```{r}
merged <- arrange(join(Activity, AvgStep_Interval), interval)
merged$steps[is.na(merged$steps)] <- merged$Avg_Steps[is.na(merged$steps)]
Activity.WithoutNA<-merged
```

Below is the histogram and summary statistics from new data set after imputing na values 

```{r}
New_Steps_perday<-aggregate(Activity.WithoutNA$steps,by=list(Activity.WithoutNA$date),FUN=sum)
names(New_Steps_perday)<-c("date","steps")

th<-theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank(),text=element_text(size=12))
p<-ggplot(New_Steps_perday,aes(x=steps,fill=..count..))+geom_histogram()
gt<-ggtitle("Histogram of steps taken each day")
myplot<-p+gt+th+ylab("Count of Steps")
myplot
summary(New_Steps_perday$steps)
```


####Are there differences in activity patterns between weekdays and weekends?

```{r}
weekdays <- weekdays(as.Date(Activity.WithoutNA$date))
data_with_weekdays <- transform(Activity.WithoutNA, day=weekdays)
data_with_weekdays$wk <- ifelse(data_with_weekdays$day %in% c("Saturday", "Sunday"),"weekend", "weekday")
average_by_interval_wk <- ddply(data_with_weekdays, .(interval, wk), summarise, steps=mean(steps))
xyplot(steps ~ interval | wk, data = average_by_interval_wk, layout = c(1, 2), type="l")
```
